---
title: "U.S. Mothers as Primary Breadwinners: Propensity and Duration"
subtitle: "Data Analysis using the SIPP 2008 data"
author: "Joanna R. Pepin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
# Getting the data
This analysis uses data downloaded from the [National Bureau of Economic Research](https://www.nber.org/data/survey-of-income-and-program-participation-sipp-data.html). We use the Core 2008 data files, waves 1-16. 

To begin, use the 0. Breadwinning_data download syntax file to set up the folder directories.  

For each wave (2008 core; 1-16):  
1. Download the zip file and then decompress the folder. Save each folder in the rawDir (generated from 0. Breadwinning_data download).  
2. Right click on the corresponding .do files and .dct files and click "save link as" and save the files in the corresponding wave folders (e.g. folder "l08puw1").

The result should be 16 folders that each contain the .dat, .dct, and .do files for each wave.

3. Change the appropriate file paths in each .do file as instructed. Delete the first line that creates a log.
``` {r, eval=F, echo=T}
*** SAMPLE STATA CODE TO REPLACE FILEPATHS
*** Asummes you have already created a "SIPP" folder and two sub-folders: "SIPP2008_raw" & "SIPP2008".
******************************************************************************************************

global datadir "C:\Users\YourName\Dropbox\Data\SIPP" /* This is your master SIPP data folder */

local dat_name "$datadir\SIPP2008_raw\l08puw1\l08puw1.dat" /*Put all wave folders into sub-folder "SIPP2008_raw" */

** The following line should contain the path to your output '.dta' file 
local dta_name "$datadir\SIPP2008\sippl08puw1.dta" /*Output all data files into sub-folder "SIPP2008" */

** The following line should contain the path to the data dictionary file 
local dct_name "$datadir\SIPP2008_raw\l08puw1\sippl08puw1.dct"

```


4. Run the .do files to create the stata data files. Use the sample code below to automate this process.  

Sample Stata code checks if each wave of the data file already exists, and runs the .do file for the wave if it doesn't exist. 
``` {r, eval=F, echo=T}
*** SAMPLE STATA CODE
*** Run this code after the file paths in each of the .do files have been changed and saved.
******************************************************************************************************

*** Change file paths as appropriate.
global datadir "C:\Users\YourName\Dropbox\Data\SIPP"
cd $datadir

*** Loop over each of the waves
forvalues w=1/16{
  capture confirm file "$datadir\SIPP2008\sippl08puw`w'.dta"
  if _rc==0 {
    display "The file File_sippl08puw`w'.dta exists"
  }
  else {
    do "$datadir/SIPP2008_raw/l08puw`w'/sippl08puw`w'"
clear all
  }
}
```

5. Import the (Stata) data files and merge them together.
``` {r, eval=T, echo=T}
library(plyr)
library(tidyverse)
library(readstata13)
library(haven)
library(data.table)

setwd(file.path(dataDir, sipp08Dir))

fileNames <- list.files(pattern = "^sippl08puw.*\\.dta$") # create a list of all waves

sipp <- lapply(fileNames, function(x) {
        read.dta13(x, 
        select.cols = c("ssuid", "shhadid", "rfid", "epppnum", "swave", "srefmon",
                        "tpearn", "thearn", "tfearn", 
                        "thothinc", "thtotinc", "tftotinc", "tpprpinc", "tptrninc", "tpothinc",
                        "tmlmsum", "amlmsum", 
                        "efnp", "t15amt", "eeducate", 
                        "aoincb1", "aoincb2", "eoincb1", "eoincb2",  
                        "tbmsum1", "tbmsum2", "abmsum1", "abmsum2",
                        "tpmsum1", "tpmsum2", "apmsum1", "apmsum2",
                        "tprftb1", "tprftb2", "aprftb1", "aprftb2",
                        "eslryb1", "eslryb2", "aslryb1", "aslryb2")) %>%
        subset(srefmon=="Fourth Reference month")
})

# Turn the list of dataframes into one master dataframe.
data <- rbindlist(sipp)

# Turn the epppnum variable into an integer
data$epppnum <- as.integer(data$epppnum)

## Subsetting data to match KR in Stata -- delete?
data <- data %>%
  subset(swave != 16)

## Save the dataframe for easy open in the future
save(data, file="sipp08tpearn_all.Rda")

```

# Creating earnings variables

Use the 1. Breadwinning_refine monthly r script

# Create Household Composition variables and merge earnings and income data to hh_change data files

Use the 2. Breadwinning_create relearn R script

This script relies on datafiles that are created using [10.1007/s13524-019-00806-1 Supplementary Materials](https://link.springer.com/article/10.1007/s13524-019-00806-1#SupplementaryMaterial). These syntax files have been updated and are available for download from this Github repository in sub-folder "childhh":

SIPP_Breadwinning
 |
 +-- childhh
 
 rmarkdown::render('breadwinning.Rmd', output_file = 'results/breadwinning.html')